\subsection{Goals} \label{subsec:goals}
% TODO support krebs Forschung

The main goal of this thesis is to identify genes that have a change in cancer activity
and are central in the extended protein-protein interaction network that we are going to create.
As these key genes are likely to be biomarkers for lung cancer,
and could be used for early detection or personalized treatment.

To achieve this goal, we have three strategic objectives:
\begin{enumerate}
    \item \textbf{Objective - Find significant changes in gene activity:}
    % Find a appropriate measurement for cancer related changes in gene activty.
    We will determine the change in gene activity between healthy and cancerous tissues.
    In order to focus on genes with significant differences later in the study.
    \label{obj:delta_tpm}

    \item \textbf{Objective - Application of a graph algorithm:}
    % We will create a graph database that contains genetic information with interactions between genes and proteins.
    We will use a graph algorithm to identify genes that are most central in the network
    to find candidates that are likely to be biomarkers for lung cancer.
    \label{obj:graph_algorithm}

    \item \textbf{Objective - Validation of results through study comparison:}
    We will compare the identified genes,
    which exhibit significant changes in activity (\ref{obj:delta_tpm})
    and are highly influential in the network (\ref{obj:graph_algorithm}) with other studies
    to validate our results and ensure that their reproducible.
    \label{obj:validation}
\end{enumerate}

% TODO Expected Outcomes
{\color{lightgray}
Our expected outcomes are:\\
A list of genes with high likelihood as potential biomarkers for lung cancer can be identified.
This gene list is confirmed by a comprehensive analysis of the literature and comparisons with existing research papers.
}