\subsection{Biological background} \label{subsec:biological_background}
To have a better understanding of our project, we need to have a closer look at some biological topics and concepts.

% Cancer overview
\textbf{Cancer} is complex, multifactorial disease characterized by uncontrolled cell growth.
Research shows that widespread metastases are the primary cause of death from cancer\cite{who_cancer_fact_sheet}.
and that cancer comes in more than 100 different types\cite{nci_cancer_types}.
The five most common forms of cancer worldwide in 2022 were lung cancer ($>2.4$ million),
breast cancer ($>2.2$ million), colorectal cancer ($>1.9$ million), prostate cancer ($>1.4$ million) and
stomach cancer ($>0.9$ million).
In 2022, cancer was one of the leading causes of death globally,
with approximately 10 million fatalities\cite{ferlay2024global}.

It can be caused by a combination of genetic and environmental factors,
such as diet, radiation, age, exposure to certain chemicals or viruses\cite{nci_cancer_risk}.
Early detection and the right treatment can significantly improve the chances of curing many types of cancer.
\\

% lung cancer
\textbf{Lung cancer} is a type of cancer that affects the lung organ.
There are two main subtypes: non-small cell carcinoma (NSCLC) and small cell carcinoma (SCLC)\cite{nci_lung_cancer_types}.
The primary risk factor for developing lung cancer is smoking,
but passive smoking and environmental pollution also significantly increase the risk\cite{nci_lung_cancer_types}.

Symptoms of lung cancer often resemble those of common colds or other minor illnesses,
such as coughing and fatigue.
This makes it challenging for patients to receive timely treatment, often leading to a diagnosis at an advanced stage\cite{who_lung_cancer}.
\\

% gene expression
\textbf{Gene expression} is a crucial factor in the investigation of cancer.
It describes the process of transcribing DNA into RNA molecules that code for proteins,
which are then translated and regulated through complex interactions to control the production and function of gene products.
\cite{nhi_gene_expression}.
Alterations in gene expression can contribute to the uncontrolled multiplication and abnormal behavior of cancer cells,
which ultimately leads to the development and progression of cancer\cite{ferlier2022regulation}.
\\

% TODO Quelle for TPM
% TPM
A quantification measure of gene expression in a cell is \textbf{Transcripts per Million (TPM)}.
A higher TPM value signifies that the corresponding gene is actively expressed and, consequently, produces more protein.
Through this method, it is possible to determine precisely which genes are activated within a cell and
their involvement in cancer development.
\\

% Gene IDs
A common way to identify genes is through the use of unique identifiers known as \textbf{ENSEMBLE IDs}.
These are unique id for genes, proteins and other genetic elements collected in the Ensemble database
from 1999 by the European Bioinformatics Institute and the Wellcome Trust Sanger Institute~\cite{ensembl_project}.
These make it easier to compare and analyze gene data across different datasets.
\\

%TODO Multi Omics Data?!